`traveltime`: an R package to calculate walking and motorised travel time from user-specified locations

### 2025-02-11 paper/paper.qmd
changes to internal referencing

---
title: "`traveltime`: an R package to calculate walking and motorised travel time from user-specified locations"
format:
  #md: default
  html:
    keep-md: true
    fig-height: 4
    fig-align: center
    fig-format: png
    dpi: 300
tags: 
  - R
  - geographic information systems
  - spatial analysis
authors:
  - name: Gerard E. Ryan
    orcid: 0000-0003-0183-7630
    corresponding: true
    affiliation: "1, 2"
  - name: Nicholas Tierney
    orcid: 0000-0003-1460-8722
    affiliation: "1, 3"
  - name: Nick Golding
    orcid: 0000-0001-8916-5570
    affiliation: "1, 4"
  - name: Daniel J. Weiss
    orcid: 0000-0002-6175-5648
    affiliation: "1, 3"
affiliations:
 - name: The Kids Research Institute Australia, Nedlands 6009 WA, Australia
   index: 1
 - name: Melbourne School of Population and Global Health, University of Melbourne, 3010, VIC, Australia
   index: 2
 - name: Curtin University, Bentley, WA, Australia
   index: 3
 - name: University of Western Australia, WA, Australia
   index: 4
date: In prep 2025
bibliography: paper.bib
---

@@

Now that we have the two items of data that we require, the next step is to prepare a friction surface for our area of interest. We will use the friction surface from @weiss2020 that can be downloaded by `traveltime` with the function `get_friction_surface()`. We can pass in our basemap `singapore_shapefile`, a `SpatVector`, directly as the `extent`. We're interested in walking time from a station, so we'll download the walking friction surface by specifying `surface = "walk2020"`. (Alternatively, we could use `surface = "motor2020"` for motorised travel if that were of interest.) We're also only interested in walking *on land* so we then mask out areas outside of the land boundary in `singapore_shapefile`:

@@

Below we plot the friction surface raster `friction_singapore`, with the vector boundary `singapore_shapefile` as a grey line, and `stations` as grey points (Figure \ref{fig-data}). `traveltime` takes resistance values of friction [@gdistance2017], so higher values of friction indicate more time travelling across a given cell. In this case walking friction is fairly uniform across the islands barring a few small inaccesible locations. More heterogeneity would be expected in less developed locations with fewer transport corridors.

@@

We present the resulting calculated travel times in Figure \ref{fig-result} where, as expected, the travel times are lowest near station exits (per Figure \ref{fig-data}) and progressively higher further away. Note that the results in `trave_time_singapore` include infinite values (`Inf`  above). In Figure \ref{fig-data}, the islands to the south and north-east are shown as filled cells, but unconnected with the mainland. The raster cells for these islands appear absent in Figure \ref{fig-result}. Because they are not connected to any cells with a station, the calculated travel time is infinite, and so these cells do not appear in Figure \ref{fig-result}.

>>>>

---
title: "`traveltime`: an R package to calculate walking and motorised travel time from user-specified locations"
format:
  jasa-pdf:
    keep-tex: true   
tags: 
  - R
  - geographic information systems
  - spatial analysis
authors:
  - name: Gerard E. Ryan
    orcid: 0000-0003-0183-7630
    corresponding: true
    affiliation: "1, 2"
  - name: Nicholas Tierney
    orcid: 0000-0003-1460-8722
    affiliation: "1, 3"
  - name: Nick Golding
    orcid: 0000-0001-8916-5570
    affiliation: "1, 4"
  - name: Daniel J. Weiss
    orcid: 0000-0002-6175-5648
    affiliation: "1, 3"
affiliations:
 - name: The Kids Research Institute Australia, Nedlands 6009 WA, Australia
   index: 1
 - name: Melbourne School of Population and Global Health, University of Melbourne, 3010, VIC, Australia
   index: 2
 - name: Curtin University, Bentley, WA, Australia
   index: 3
 - name: University of Western Australia, WA, Australia
   index: 4
date: In prep 2025
bibliography: paper.bib
---

@@

Now that we have the two items of data that we require, the next step is to prepare a friction surface for our area of interest. We will use the friction surface from @weiss2020global that can be downloaded by `traveltime` with the function `get_friction_surface()`. We can pass in our basemap `singapore_shapefile`, a `SpatVector`, directly as the `extent`. We're interested in walking time from a station, so we'll download the walking friction surface by specifying `surface = "walk2020"`. (Alternatively, we could use `surface = "motor2020"` for motorised travel if that were of interest.) We're also only interested in walking *on land* so we then mask out areas outside of the land boundary in `singapore_shapefile`:

@@

Below we plot the friction surface raster `friction_singapore`, with the vector boundary `singapore_shapefile` as a grey line, and `stations` as grey points (Figure @fig-data). `traveltime` takes resistance values of friction [@gdistance2017], so higher values of friction indicate more time travelling across a given cell. In this case walking friction is fairly uniform across the islands barring a few small inaccesible locations. More heterogeneity would be expected in less developed locations with fewer transport corridors.

@@

We present the resulting calculated travel times in Figure @fig-result where, as expected, the travel times are lowest near station exits (per Figure @fig-data) and progressively higher further away. Note that the results in `trave_time_singapore` include infinite values (`Inf`  above). In Figure @fig-data, the islands to the south and north-east are shown as filled cells, but unconnected with the mainland. The raster cells for these islands appear absent in Figure @fig-result. Because they are not connected to any cells with a station, the calculated travel time is infinite, and so these cells do not appear in Figure @fig-result.
